Occultation Predictions of Irregular Satellites

### 2016-04-15 relatorio/Occ-Irregular/Relatorio.tex
.

% mn2esample.tex
%
% v2.1 released 22nd May 2002 (G. Hutton)
%
% The mnsample.tex file has been amended to highlight
% the proper use of LaTeX2e code with the class file
% and using natbib cross-referencing. These changes
% do not reflect the original paper by A. V. Raveendran.
%
% Previous versions of this sample document were
% compatible with the LaTeX 2.09 style file mn.sty
% v1.2 released 5th September 1994 (M. Reed)
% v1.1 released 18th July 1994
% v1.0 released 28th January 1994

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage[brazilian,english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{prettyref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{epstopdf}
\usepackage{color}
\usepackage[breaklinks=true]{hyperref}  
\usepackage{arydshln} % creates horizontal dashed lines in tables

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%

\title[New orbits and predictions of stellar occultations by irregular satellites up to 2020]{New orbits for irregular satellites based on thousands of new observations and predictions of stellar occultations up to 2020}
\author[A. R. Gomes-J\'unior, M. Assafin, L. Beauvalet et al.]{A. R. Gomes-J\'unior$^{1}$\thanks{E-mail: altair08@astro.ufrj.br},
M. Assafin$^{1,\dag}$,
L. Beauvalet$^{2,3}$,
J. Desmars$^{4}$,
R. Vieira-Martins$^{1,2,5,\dag}$,\newauthor
J. I. B. Camargo$^{2,5}$,
B. E. Morgado$^{1,2}$
F. Braga-Ribas$^{2,6}$
\\
$^{1}$Observat\'orio do Valongo/UFRJ, Ladeira Pedro Ant\^onio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
$^{2}$Observat\'orio Nacional/MCTI, R. General Jos\'e Cristino 77, CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
$^{3}$Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau 75014 Paris, France\\
$^{4}$Institut de m\'ecanique c\'eleste et de calcul des \'eph\'em\'erides - Observatoire de Paris, UMR 8028 du CNRS, 77 Av. Denfert-Rochereau,\\ 75014 Paris, France\\
$^{5}$Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ 20921-400, Brazil\\
$^{6}$Federal University of Technology - Paran\'a (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil
}
\begin{document}
\newcommand{\noccs}{5442 } % total number of occultations

\date{Accepted . Received ; in original form }

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2016}

@@

\label{firstpage}

\begin{abstract}
%Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, shape, albedo and composition, to trace back their true origin. The best ground-based technique to determine size and shape, and thus constrain the albedo and in a broader sense composition, is the observation of stellar occultations by these objects but this demands a very precise ephemeris based on new and good positions.
%\cite{GomesJunior2015} published 6523 observations obtained between 1992 and 2014 for 18 irregular satellites with an estimated error in the positions of about 60 to 80 mas.
%Based on this observations, we aimed to generate new numerical integration of the orbits of eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda. This is a short-time ephemeris that corrected systematic errors from previous ephemeris (JPL JUP300). For Phoebe we updated ephemeris from \cite{Desmars2013} using 75\% more positions than the previous one.
%Also Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018. This creates a good opportunity to predict stellar occultations due to the large number of events that can be predicted in these periods. Thus, we aimed to predict stellar occultations for the same 9 satellites cited above. Using the UCAC4 catalogue we managed to identify \noccs candidates of stellar occultations between the period of January, 2016 and December, 2020.
%We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.
\cite{GomesJunior2015} published 6523 new positions for 18 irregular satellites from observations made between 1992 and 2014 with an estimated error of about 60 to 80 mas. Based on this set of original positons, we performed numerical integrations for deriving new orbits for the eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda. This corrected systematic errors and resulted in an ephemeris more accurate/precise than JPL JUP300 for these bodies at least in the immediate future. For Phoebe we updated the ephemeris from \cite{Desmars2013} using 75\% more positions than the previous one. Due to their orbital configurations, it is common belief that the irregular satellites were captured by the giant planets in the early Solar System, but there is no consensus for a single model explaining where they were formed. Size, shape, albedo and composition would help a lot to trace back their true origin, but these physical parameters are yet poorly known for irregular satellites. The observation of stellar occultations would allow for the determination of such parameters. Indeed Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018, improving a lot the chances of observing such events in the near future. Using the derived ephemerides and the UCAC4 catalogue we managed to identify \noccs candidate stellar occultations between January 2016 and December 2020 for the 9 irregular satellites studied here. We discussed how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.
\end{abstract}

\begin{keywords}
occultations - planets and satellites: general - planets and satellites: individual: Jovian and Saturnian irregular satellites - planets and satellites: individual: Triton
\end{keywords}

\section{Introduction}\label{Sec: introducao}

@@

\cite{GomesJunior2015} obtained 6523 suitable positions for 18 irregular satellites between 1992 and 2014 with an estimated error in the positions of about 60 to 80 mas. For some satellites the number of positions obtained is comparable to the number used in the numerical integration of orbits by the JPL \citep{Jacobson2012}. They pointed out that the ephemeris of the irregular satellites have systematic errors that may reach 200 mas for some satellites.

We present in this paper new numerical integration of the orbits of the 8 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke, Sinope and Leda) using only the positions obtained by \cite{GomesJunior2015}. We aim to eliminate the systematic errors in the JPL ephemeris (see Sec. \ref{Sec: integration}). %For an object at the distance of Jupiter, this represents an error langer than 700 km in the shadow path of an occultation.
For Phoebe, we updated the ephemeris of \cite{Desmars2013} using the observations of \cite{GomesJunior2015}, \cite{Peng2015}, observations from Minor Planet Center and observations from Flagstaff.

@@

In section \ref{Sec: integration} the new determination of the orbits. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites, including some tests made to check the accuracy of the predictions. The final discussion is presented in section \ref{Sec: discussion}.

\section{Orbit computations} \label{Sec: integration}

@@

\subsection{Special tailored Ephemerides (STE) for Jupiter irregular satellites}

@@

Our numerical model describes the dynamical evolution of the irregular satellites of Jupiter in a jovicentric reference frame. The satellites are submitted to the influence of the Sun and the rest of the solar system, as well as that of the Galilean satellites and the first harmonics of Jupiter's gravity field. The axes of the reference frame are those of the ICRS.

We use the following notations:
\begin{itemize}
\item $i$ and $l$ one of the irregular satellites of Jupiter
\item $J$ Jupiter 
\item $j$ another body of the Solar System
\item $M_j$ the mass of the $j$-th body, not an irregular satellite
\item $m_i$ the mass of the irregular satellite $i$
\item $\vec{r_i}$ the position of the $i$-th body with respect to the barycentre of Jupiter System
\item $r_{ij}$ the distance between bodies $i$ and $j$  
\item $R_J$ the radius of Jupiter
\item $J_n$ the dynamic polar oblateness of the n-th order for Jupiter's gravity field
\item $U_{\bar{l}\hat{J}}$ potential generated by the oblateness of Jupiter on the satellite $l$
\item $\Phi_i$ is the inclination of the $i$-th satellite with respect to Jupiter's equator.
\end{itemize}

For an irregular satellite $i$, under the gravitational influence of Jupiter, the $\mathcal{N}-1$ other irregular satellites, the regular Jovian satellites and the rest of the Solar System ($\mathcal{N}$ bodies), the equation of motion is:
\begin{equation}\begin{array}{ll}

\ddot{\vec{r_i}}= & \displaystyle -GM_J\frac{\vec{r_J}-\vec{r_i}}{r_{iJ}^3}-\sum_{l=1,l\neq i}^\mathcal{N'}Gm_l\frac{\vec{r_l}-\vec{r_i}}{r_{il}^3}\\
&\displaystyle -\sum_{j=1}^\mathcal{N}GM_j \left(\frac{\vec{r_j}-\vec{r_i}}{r_{ij}^3} - \frac{\vec{r_j}-\vec{r_J}}{r_{Jj}^3} \right)\\
 & \displaystyle +GM_J \nabla U_{\bar{l}\hat{J}} -\sum_{l=1}^\mathcal{N} Gm_l\nabla U_{\bar{l}\hat{J}}
\end{array}
\label{Eq:1}
\end{equation}
where the last term in the brackets and the last term in Eq. \ref{Eq:1} represent undirect perturbations. The oblateness potential seen by the body $i$ because of Jupiter is:
\begin{equation}\begin{array}{ll}

U_{\bar{l}\hat{J}}=&\displaystyle -\frac{R_J^2 J_2}{r_{iJ}^3}\left(\frac{3}{2}\sin^2 \Phi_i-\frac{1}{2}\right)\\ &\\ & 
\displaystyle-\frac{R_J^4 J_4}{r_{iJ}^5}\left(\frac{35}{8}\sin^4 \Phi_i-\frac{15}{4}\sin^2 \Phi_i+\frac{3}{8}\right)\\
& \\
&\displaystyle-\frac{R_J^6 J_6}{r_{iJ}^7}\left(\frac{231}{16}\sin^6 \Phi_i-\frac{315}{16}\sin^4 \Phi_i+\frac{105}{16}\sin^2 \Phi_i-\frac{5}{16}\right)

\end{array}
\end{equation}

The expressions of $\nabla_lU_{\bar{l}\, \hat{i}}$ and $\nabla_i U_{\bar{i}\, \hat{l}}.$ have been developed in \citet{Lainey2004}. The equations of motion are integrated with the numerical integrator RADAU \citep{Everhart1985}. 
Our model was fitted to the observations through a least-squares procedure. The satellites were integrated one dynamical family at a time, to gain computing time, while losing minimum precision. Indeed, the interactions between satellites not belonging to the same dynamical family are negligible considering the short timespan of our integration.

The initial osculating elements at the origin of integration are presented in Table \ref{Tab: sat_ell}.

\begin{table*}
\caption{Initial osculating elements for Jupiter irregular satellites at JD 2451545.0.
 }\label{Tab: sat_ell}
\begin{center}
\begin{tabular}{ccccccc}
\hline\hline
Satellite & a (km) & e & I$\degr$ & $\Omega\degr$ & $\omega\degr$ & $v\degr$ \\ 
\hline
Himalia &   11372100 $\pm$ 500    &    0.166 $\pm$ 0.002      &   45.14 $\pm$ 0.15      &   39.77 $\pm$ 0.19      &   351.48 $\pm$ 0.46      &   97.35 $\pm$ 0.48    \\
Elara &   11741170 $\pm$ 690  &      0.222 $\pm$ 0.002      &   28.64 $\pm$ 0.18      &   68.42 $\pm$ 0.43      &   179.82 $\pm$ 0.56      &   339.08 $\pm$ 0.82  \\
Lysithea &   11739900 $\pm$ 1300  &      0.136 $\pm$ 0.004      &    51.12 $\pm$ 0.27     &   5.53 $\pm$ 0.52      &   53.0 $\pm$ 1.5      &   318.9 $\pm$ 2.0   \\
Leda &   11140300  $\pm$ 4300  &     0.173  $\pm$ 0.007     &   16.15  $\pm$ 0.75    &   272.6  $\pm$ 1.7    &   212.2  $\pm$ 3.6          &   218.8  $\pm$ 3.2  \\
Pasiphae &  23425000  $\pm$ 5000    &     0.379  $\pm$ 0.001       &   152.44 $\pm$ 0.10      &   284.59 $\pm$ 0.21      &   135.96 $\pm$ 0.19      &   236.97 $\pm$ 0.16 \\
Sinope &   22968800 $\pm$ 5200   &     0.316 $\pm$ 0.002      &   157.76 $\pm$ 0.12      &   256.62 $\pm$ 0.55      &   298.38 $\pm$ 0.55      &   167.57 $\pm$ 0.19    \\
Carme &   24202924 $\pm$ 4800      &  0.242 $\pm$ 0.001      &   147.13 $\pm$ 0.10      &   154.01 $\pm$ 0.25      &   47.90 $\pm$ 0.29      &   234.41 $\pm$ 0.19  \\
Ananke &  21683800  $\pm$ 7200  &     0.380 $\pm$ 0.002      &   172.29 $\pm$ 0.20      &   56.9 $\pm$ 1.2      &   123.3 $\pm$ 1.2      &   231.24 $\pm$ 0.21  \\
%Themisto  & 7393800    &     0.198       &   25.77  &   220.0     &   216.3    &   262.1     \\
\hline
\end{tabular} 
\end{center}
\begin{flushleft}
\textbf{Notes}: a: semimajor axis; e : excentricity; I: inclination relative to the equatorial reference plane J2000; $\Omega$: longitude of the ascending node; $\omega$: argument of periapsis; $v$: true anomaly.
\end{flushleft}
\end{table*}

@@

\begin{figure}
\includegraphics[width=8.8cm]{figures/Carme_ephemeris.eps}
\caption{Offsets in declination of the positions published by \protect\cite{GomesJunior2015} for Carme. The red "x" relative to the special-tailored ephemeris, the blue "+" relative to the JUP300 JPL ephemeris and the green dot relative to \protect\cite{Emelyanov2008}. As expected, the ephemeris systematic errors pointed out by \protect\cite{GomesJunior2015} are reduced with the STE ephemeris. \label{Fig: JPL-STE}}
\end{figure}

@@

\subsection{Phoebe's ephemeris}

@@

\begin{figure}
\begin{centering}
\includegraphics[width=8.8cm]{figures/Phoebe.eps} 
\caption{Comparison between the PH15, SAT375 JPL and \protect\cite{Emelyanov2007} ephemeris for the satellite Phoebe.}
\label{Fig:eph-Phoebe}
\end{centering}
\end{figure}

@@

\section{Prediction of stellar occultations} \label{Sec: predictions}

\subsection{Candidate events}

The prediction of the occultations was made by crossing the stellar coordinates and proper motions of the UCAC4 catalogue \citep{Zacharias2013} with the ephemeris presented in Sec. \ref{Sec: integration}. The search for stellar candidates follows the same procedure as presented by \cite{Assafin2010, Assafin2012} and \cite{Camargo2014}.

@@

\begin{table}
\caption{\label{Tab: satellite-occultation} Number of stellar occultations for each satellite from January, 2016 up to December, 2020.}
\begin{centering}
\begin{tabular}{lcccccc}
\hline  \hline
%\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & 2016 & 2017 & 2018 & 2019 & 2020 & Total \tabularnewline
\hline
Ananke & 12 & 16 & 49 & 359 & 187 & 623 \tabularnewline
Carme & 20 & 14 & 30 & 369 & 220 & 653 \tabularnewline
Elara & 14 & 16 & 33 & 305 & 193 & 561 \tabularnewline
Himalia & 15 & 12 & 54 & 257 & 230 & 568 \tabularnewline
Leda & 8 & 24 & 38 & 362 & 208 & 640 \tabularnewline
Lysithea & 16 & 11 & 35 & 330 & 212 & 604 \tabularnewline
Pasiphae & 20 & 19 & 44 & 362 & 206 & 651 \tabularnewline
Sinope & 15 & 21 & 34 & 356 & 256 & 682 \tabularnewline
\hdashline
Phoebe & 32 & 98 & 238 & 79 & 13 & 460 \tabularnewline
%\hdashline
%Nereid & 11$^{a}$ & 1 & 1 & 0 & 0 & 13 \tabularnewline
%Triton & -- & 1 & 1 & 0 & 0 & 2 \tabularnewline
\hline
\end{tabular}
\par \end{centering}
$^{a}$ {Using the WFI catalogue as explained in Sec. \ref{Sec: predictions}}.
\end{table}

Table \ref{Tab: sample-cds} shows a sample of the catalogue of occultations generated and their parameters, which are necessary to produce occultation maps. Since these objects are very small, the duration of each event is a few seconds. All the occultation tables and maps will be publicly available at the CDS. In Fig. \ref{Fig: ocultacao} we show an example of an occultation map. This is an occultation by Elara that will happen in February 21, 2017. This event can be observed from Australia and it is one of the best opportunities for this object due to the slow velocity of the event and it involves a bright star (MagR*=12.4).

>>>>

\documentclass[12pt,a4paper]{report}
\usepackage[brazilian, english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsthm,amsfonts,amssymb,textcomp}
%\usepackage{latexsym}
\usepackage{graphicx}
\graphicspath{{figuras}}
\usepackage{subfigure}
\usepackage{float}
\usepackage{longtable}
\usepackage{color}
\usepackage{epstopdf}
\usepackage{pdflscape}
\usepackage[breaklinks=true]{hyperref}
\usepackage[comma,authoryear]{natbib}
\usepackage[nonumberlist]{glossaries}
\usepackage{arydshln}
\usepackage{footnote}
\usepackage{longtable}
\usepackage[small,bf,singlelinecheck=off]{caption}
\usepackage[left=3cm,right=2cm,top=3cm,bottom=2cm]{geometry}
\usepackage{arydshln}
\usepackage{scalefnt}
%\usepackage[alf]{abntcite}

%%% newcommand %%%%%%%%%%%%%%%%%
\newcommand{\PE}{Perkin-Elmer }
\newcommand{\BC}{Boller \& Chivens }
\newcommand{\noccs}{5442 }

\newcommand{\degr}{\ensuremath{^{\circ}}}%                    % degree symbol:  °
\newcommand{\arcmin}{\ensuremath{^{\prime}}}%                    % degree symbol:  °
\newcommand{\arcsec}{\ensuremath{^{\prime\prime}}}%                    % degree symbol:  °
\newcommand{\fs}{\mbox{\ensuremath{.\!\!^s}}}
\newcommand{\farcm}{\mbox{\ensuremath{.\mkern-4mu^\prime}}}%    % fractional arcminute symbol: 0.'0
\newcommand{\farcs}{\mbox{\ensuremath{.\!\!^{\prime\prime}}}}%  % fractional arcsecond symbol: 0.''0
\newcommand{\fdg}{\mbox{\ensuremath{.\!\!^\circ}}}%             % fractional degree symbol:     0.°0

%\makeglossaries

\makeatletter
\renewcommand\chapter{\thispagestyle{plain}
                \global\@topnum\z@
                \@afterindentfalse
                \secdef\@chapter\@schapter}
\makeatother

\makeatletter
\renewcommand{\@makechapterhead}[1]{%
\vspace*{50 pt}%
{\setlength{\parindent}{0pt} \raggedright \normalfont
\bfseries\Huge\thechapter.\ #1
\par\nobreak\vspace{40 pt}}}
\makeatother

%\input{glossario}
\author{Altair Ramos Gomes Júnior}
\title{Occultation Predictions of Irregular Satellites}
\begin{document}

@@

\pagestyle{headings}

\section*{Introduction}\label{Sec: introducao}

@@

Size, shape, albedo and composition would help a lot to trace back their true origin, but these physical parameters are yet poorly known for irregular satellites. The observation of stellar occultations would allow for the determination of such parameters. Indeed Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018, improving a lot the chances of observing such events in the near future.

%\cite{GomesJunior2015} obtained 6523 suitable positions for 18 irregular satellites between 1992 and 2014 with an estimated error in the positions of about 60 to 80 mas. For some satellites the number of positions obtained is comparable to the number used in the numerical integration of orbits by the JPL \citep{Jacobson2012}. They pointed out that the ephemeris of the irregular satellites have systematic errors that may reach 200 mas for some satellites.

For this, we made new numerical integration of the orbits of the 8 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke, Sinope and Leda) using only the positions obtained by \cite{GomesJunior2015}. We aim to eliminate the systematic errors in the JPL ephemeris pointed out by these authors. %For an object at the distance of Jupiter, this represents an error langer than 700 km in the shadow path of an occultation.
For Phoebe, we updated the ephemeris of \cite{Desmars2013} using the observations of \cite{GomesJunior2015}, \cite{Peng2015}, observations from Minor Planet Center and observations from Flagstaff. With this, we predicted stellar occultation for the 9 objects.

Using the derived ephemerides and the UCAC4 catalogue we managed to identify \noccs candidate stellar occultations between January 2016 and December 2020 for the 9 irregular satellites studied here.

@@

\section*{Orbit computations} \label{Sec: integration}

@@

\subsection*{Special tailored Ephemerides (STE) for Jupiter irregular satellites}

@@

%Our numerical model describes the dynamical evolution of the irregular satellites of Jupiter in a jovicentric reference frame. The satellites are submitted to the influence of the Sun and the rest of the solar system, as well as that of the Galilean satellites and the first harmonics of Jupiter's gravity field. The axes of the reference frame are those of the ICRS. 
%
%We use the following notations:
%\begin{itemize}
%\item $i$ and $l$ one of the irregular satellites of Jupiter
%\item $J$ Jupiter 
%\item $j$ another body of the Solar System
%\item $M_j$ the mass of the $j$-th body, not an irregular satellite
%\item $m_i$ the mass of the irregular satellite $i$
%\item $\vec{r_i}$ the position of the $i$-th body with respect to the barycentre of Jupiter System
%\item $r_{ij}$ the distance between bodies $i$ and $j$  
%\item $R_J$ the radius of Jupiter
%\item $J_n$ the dynamic polar oblateness of the n-th order for Jupiter's gravity field
%\item $U_{\bar{l}\hat{J}}$ potential generated by the oblateness of Jupiter on the satellite $l$
%\item $\Phi_i$ is the inclination of the $i$-th satellite with respect to Jupiter's equator.
%\end{itemize}
%
%For an irregular satellite $i$, under the gravitational influence of Jupiter, the $\mathcal{N}-1$ other irregular satellites, the regular Jovian satellites and the rest of the Solar System ($\mathcal{N}$ bodies), the equation of motion is:
%\begin{equation}\begin{array}{ll}
%
%\ddot{\vec{r_i}}= & \displaystyle -GM_J\frac{\vec{r_J}-\vec{r_i}}{r_{iJ}^3}-\sum_{l=1,l\neq i}^\mathcal{N'}Gm_l\frac{\vec{r_l}-\vec{r_i}}{r_{il}^3}\\
%&\displaystyle -\sum_{j=1}^\mathcal{N}GM_j \left(\frac{\vec{r_j}-\vec{r_i}}{r_{ij}^3} - \frac{\vec{r_j}-\vec{r_J}}{r_{Jj}^3} \right)\\
% & \displaystyle +GM_J \nabla U_{\bar{l}\hat{J}} -\sum_{l=1}^\mathcal{N} Gm_l\nabla U_{\bar{l}\hat{J}}
%\end{array}
%\label{Eq:1}
%\end{equation}
%where the last term in the brackets and the last term in Eq. \ref{Eq:1} represent undirect perturbations. The oblateness potential seen by the body $i$ because of Jupiter is:
%\begin{equation}\begin{array}{ll}
%
%U_{\bar{l}\hat{J}}=&\displaystyle -\frac{R_J^2 J_2}{r_{iJ}^3}\left(\frac{3}{2}\sin^2 \Phi_i-\frac{1}{2}\right)\\ &\\ & 
%\displaystyle-\frac{R_J^4 J_4}{r_{iJ}^5}\left(\frac{35}{8}\sin^4 \Phi_i-\frac{15}{4}\sin^2 \Phi_i+\frac{3}{8}\right)\\
%& \\
%&\displaystyle-\frac{R_J^6 J_6}{r_{iJ}^7}\left(\frac{231}{16}\sin^6 \Phi_i-\frac{315}{16}\sin^4 \Phi_i+\frac{105}{16}\sin^2 \Phi_i-\frac{5}{16}\right)
%
%\end{array}
%\end{equation}
%
%The expressions of $\nabla_lU_{\bar{l}\, \hat{i}}$ and $\nabla_i U_{\bar{i}\, \hat{l}}.$ have been developed in \citet{Lainey2004}. The equations of motion are integrated with the numerical integrator RADAU \citep{Everhart1985}. 
%Our model was fitted to the observations through a least-squares procedure. The satellites were integrated one dynamical family at a time, to gain computing time, while losing minimum precision. Indeed, the interactions between satellites not belonging to the same dynamical family are negligible considering the short timespan of our integration. 
%
%The initial osculating elements at the origin of integration are presented in Table \ref{Tab: sat_ell}.
%
%\begin{table*}
%\scalefont{0.8}
%\caption{Initial osculating elements for Jupiter irregular satellites at JD 2451545.0.
% }\label{Tab: sat_ell}
%\begin{center}
%\begin{tabular}{ccccccc}
%\hline\hline
%Satellite & a (km) & e & I$\degr$ & $\Omega\degr$ & $\omega\degr$ & $v\degr$ \\ 
%\hline
%Himalia &   11372100 $\pm$ 500    &    0.166 $\pm$ 0.002      &   45.14 $\pm$ 0.15      &   39.77 $\pm$ 0.19      &   351.48 $\pm$ 0.46      &   97.35 $\pm$ 0.48    \\
%Elara &   11741170 $\pm$ 690  &      0.222 $\pm$ 0.002      &   28.64 $\pm$ 0.18      &   68.42 $\pm$ 0.43      &   179.82 $\pm$ 0.56      &   339.08 $\pm$ 0.82  \\
%Lysithea &   11739900 $\pm$ 1300  &      0.136 $\pm$ 0.004      &    51.12 $\pm$ 0.27     &   5.53 $\pm$ 0.52      &   53.0 $\pm$ 1.5      &   318.9 $\pm$ 2.0   \\
%Leda &   11140300  $\pm$ 4300  &     0.173  $\pm$ 0.007     &   16.15  $\pm$ 0.75    &   272.6  $\pm$ 1.7    &   212.2  $\pm$ 3.6          &   218.8  $\pm$ 3.2  \\
%Pasiphae &  23425000  $\pm$ 5000    &     0.379  $\pm$ 0.001       &   152.44 $\pm$ 0.10      &   284.59 $\pm$ 0.21      &   135.96 $\pm$ 0.19      &   236.97 $\pm$ 0.16 \\
%Sinope &   22968800 $\pm$ 5200   &     0.316 $\pm$ 0.002      &   157.76 $\pm$ 0.12      &   256.62 $\pm$ 0.55      &   298.38 $\pm$ 0.55      &   167.57 $\pm$ 0.19    \\
%Carme &   24202924 $\pm$ 4800      &  0.242 $\pm$ 0.001      &   147.13 $\pm$ 0.10      &   154.01 $\pm$ 0.25      &   47.90 $\pm$ 0.29      &   234.41 $\pm$ 0.19  \\
%Ananke &  21683800  $\pm$ 7200  &     0.380 $\pm$ 0.002      &   172.29 $\pm$ 0.20      &   56.9 $\pm$ 1.2      &   123.3 $\pm$ 1.2      &   231.24 $\pm$ 0.21  \\
%%Themisto  & 7393800    &     0.198       &   25.77  &   220.0     &   216.3    &   262.1     \\
%\hline
%\end{tabular} 
%\end{center}
%\begin{flushleft}
%\textbf{Notes}: a: semimajor axis; e : excentricity; I: inclination relative to the equatorial reference plane J2000; $\Omega$: longitude of the ascending node; $\omega$: argument of periapsis; $v$: true anomaly.
%\end{flushleft}
%\end{table*}

@@

\begin{figure}
\includegraphics[width=16.0cm]{figures/Carme_ephemeris.eps}
\caption{Offsets in declination of the positions published by \protect\cite{GomesJunior2015} for Carme. The red "x" relative to the special-tailored ephemeris, the blue "+" relative to the JUP300 JPL ephemeris and the green dot relative to \protect\cite{Emelyanov2008}. As expected, the ephemeris systematic errors pointed out by \protect\cite{GomesJunior2015} are reduced with the STE ephemeris. \label{Fig: JPL-STE}}
\end{figure}

@@

\subsection*{Phoebe's ephemeris}

@@

\begin{figure}
\begin{centering}
\includegraphics[width=16.0cm]{figures/Phoebe.eps} 
\caption{Comparison between the PH15, SAT375 JPL and \protect\cite{Emelyanov2007} ephemeris for the satellite Phoebe.}
\label{Fig:eph-Phoebe}
\end{centering}
\end{figure}

@@

\section*{Prediction of stellar occultations} \label{Sec: predictions}

\subsection*{Candidate events}

The prediction of the occultations was made by crossing the stellar coordinates and proper motions of the UCAC4 catalogue \citep{Zacharias2013} with the ephemeris presented in the previous section. The search for stellar candidates follows the same procedure as presented by \cite{Assafin2010, Assafin2012} and \cite{Camargo2014}.

@@

\begin{table}[h]
\caption{\label{Tab: satellite-occultation} Number of stellar occultations for each satellite from January, 2016 up to December, 2020.}
\begin{centering}
\begin{tabular}{lcccccc}
\hline  \hline
%\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & 2016 & 2017 & 2018 & 2019 & 2020 & Total \tabularnewline
\hline
Ananke & 12 & 16 & 49 & 359 & 187 & 623 \tabularnewline
Carme & 20 & 14 & 30 & 369 & 220 & 653 \tabularnewline
Elara & 14 & 16 & 33 & 305 & 193 & 561 \tabularnewline
Himalia & 15 & 12 & 54 & 257 & 230 & 568 \tabularnewline
Leda & 8 & 24 & 38 & 362 & 208 & 640 \tabularnewline
Lysithea & 16 & 11 & 35 & 330 & 212 & 604 \tabularnewline
Pasiphae & 20 & 19 & 44 & 362 & 206 & 651 \tabularnewline
Sinope & 15 & 21 & 34 & 356 & 256 & 682 \tabularnewline
\hdashline
Phoebe & 32 & 98 & 238 & 79 & 13 & 460 \tabularnewline
%\hdashline
%Nereid & 11$^{a}$ & 1 & 1 & 0 & 0 & 13 \tabularnewline
%Triton & -- & 1 & 1 & 0 & 0 & 2 \tabularnewline
\hline
\end{tabular}
\par \end{centering}
\end{table}

%Table \ref{Tab: sample-cds} shows a sample of the catalogue of occultations generated and their parameters, which are necessary to produce occultation maps. Since these objects are very small, the duration of each event is a few seconds. All the occultation tables and maps will be publicly available at the CDS.
In Fig. \ref{Fig: ocultacao} we show an example of an occultation map. This is an occultation by Elara that will happen in February 21, 2017. This event can be observed from Australia and it is one of the best opportunities for this object due to the slow velocity of the event and it involves a bright star (MagR*=12.4).

%\begin{table*}
%\caption{\label{Tab: sample-cds} A sample of stellar occultation predictions for Pasiphae}
%\begin{center}
%\begin{tabular}{cccrcccrcrr}
%\hline
%\hline
%d m Year~~~~h~~m~~~s & RA~~~(ICRS)~~~Dec & C/A & P/A & $\nu$ & {\it D} & $R^*$ & $\lambda$ & LST & $\mu_{\alpha*}$ & $\mu_{\delta}$ \tabularnewline
%\hline
%% & \multicolumn{2}{c}{Mean errors} & Nr & Nr &  UCAC4       \tabularnewline
%09 04 2016 03:58:19. & 11 14 36.7707 +07 39 20.7610 & 1.003 &  17.9 & -12.88 &  4.54 & 14.9 & 271. & 22:03 &  12. & -33. \tabularnewline 
%13 06 2016 00:16:12. & 11 12 48.5020 +07 06 43.3520 & 0.661 &  30.0 & +14.32 &  5.50 & 13.9 & 262. & 17:45 &  -1. &   1. \tabularnewline 
%27 06 2016 13:56:09. & 11 18 03.4160 +06 23 45.1940 & 1.707 &  28.0 & +20.29 &  5.74 & 11.7 &  44. & 16:53 &   4. & -10. \tabularnewline 
%18 07 2016 15:07:24. & 11 28 15.5076 +05 05 31.8060 & 0.942 &  26.7 & +27.80 &  6.05 & 14.0 &   8. & 15:40 &   4. &   4. \tabularnewline 
%22 07 2016 16:15:07. & 11 30 30.4310 +04 48 43.4340 & 0.644 & 206.5 & +29.04 &  6.11 & 14.6 & 348. & 15:27 &  23. & -24. \tabularnewline 
%24 07 2016 01:37:34. & 11 31 17.8471 +04 42 49.0540 & 0.029 & 206.6 & +29.46 &  6.12 & 15.1 & 206. & 15:22 &   2. &  -8. \tabularnewline 
%24 07 2016 17:37:18. & 11 31 40.7472 +04 39 57.5060 & 0.840 &  26.5 & +29.66 &  6.13 & 14.9 & 326. & 15:20 & -11. &  -1. \tabularnewline 
%\hline
%\end{tabular}
%\end{center}
%\begin{flushleft}
%\textbf{Notes}. Entries included: day of the year and UTC time of the prediction; right ascension and declination of the occulted star - at the central instant of occultation (corrected by proper motions); C/A: the geocentric closest approach, in arcseconds; P/A: the satellite position angle with respect to the occulted star at C/A, in degrees (zero at north of the star, increasing clockwise); $\nu$: velocity in the plane of sky, in km s$^{-1}$: positive = prograde, negative = retrograde; {\it D}: planet range to Earth, in AU; $R^*$: normalized magnitude to a common shadow velocity of 20 km s$^{-1}$ by the relationship $\textrm{R}^* = \textrm{R}_{\textrm{actual}} + 2.5 \times \log 10 \left(\frac{\textrm{velocity}} {20 \textrm{km}\, \textrm{s}^{-1}} \right)$; $\lambda$: east longitude of subplanet point in degrees, positive towards east; LST: UT + $\lambda$: local solar time at subplanet point, hh:mm; $\mu_{\alpha *}$ and $\mu_{\delta}$: proper motions in right ascension and declination, respectively (mas/year).
%\end{flushleft}
%\end{table*}

\begin{figure}
\begin{centering}
\includegraphics[width = 16.0cm]{figures/Elara_2017-02-21T15:28:40.png}   
\caption{Occultation map for Elara. % regarding to the first event sampled in Table \ref{Tab: occ-list}.
The central red dot show the geocentric closest approach of the shadow. The small ones shows the center of the shadow separated by 60s. The lines show the path of the shadow over the Earth. The shadow moves from right to left.
\textbf{Labels:} Diam: Diameter of the object; Tmax: Maximum duration of the event for a central observation; C/A: the geocentric closest approach, in arcseconds; P/A: the satellite position angle with respect to the occulted star at C/A, in degrees; vel: velocity of event in km/s; Delta: Geocentric distance to the occulting object in AU; $R^*$: normalized magnitude to a common shadow velocity of 20 km s$^{-1}$; long: east longitude of subplanet point in degrees, positive towards east.}
\label{Fig: ocultacao}
\end{centering}
\end{figure}
